A Study on Common Bug Fix Patterns in Rust

### 2022-09-20 sections/methodology.tex
reorganized files
warnings fixed

An ASTDiff may include an arbitrary number of semantic changes. We are interested in finding the most important change within an ASTDiff and encoding that change in our datapoints. Sections ? and ? provide a detailed explanation of how we select the most important semantic information, which then helped us to obtain clusters that contained similar datapoints.

We used Pydriller, which is a repository miner for Python. Our target repositories were the top 18 most starred open source Rust projects on GitHub. We mined all the bug related commits and ran them through our pipeline. Then, we applied the DBSCAN clustering algorithm on the obtained datapoints. In Section ? we discuss why we chose DBSCAN for clustering and how we tuned it to improve clusters quality.

\subsection{Data modelling}

\subsubsection{Parsing Programs}

@@

\subsubsection{Path Extraction}

\begin{figure}[h]
\centering
\includegraphics[width=0.8\textwidth]{extraction.png}
\caption{\label{fig:extraction}Path Extraction}
\end{figure}

@@

\subsubsection{Weighting Scheme}

@@

\subsection{Mining Repositories}

\begin{figure}[h]
\centering
\includegraphics[width=1\textwidth]{repos.png}
\caption{\label{table:repos} Target Repositories}
\end{figure}

@@

\subsection{Clustering Data}

@@

\begin{figure}[h]
\centering
\includegraphics[width=1\textwidth]{clusters.png}
\caption{\label{fig:clustering} Clustering Experiments}
\end{figure}

@@

\subsubsection{Manual Analysis for parameter tuning}

@@

\subsubsection{Manual Analysis for cluster selection}
% 3 types, refactoring dropped

@@

In our manual analysis, we randomly picked 50 datapoints of each cluster. If the cluster had less than 50 datapoints, we would analyze all of the datapoints in the cluster. Next, we read the code and also the bug report related to all the datapoints (if existed). In case that a similar pattern was detected over the instances of a cluster, we would write a natural language description of that cluster. After providing all the descriptions, we linked similar clusters together as possible candidates for merging. At last, after accomplishing re-examination for merge possibilities, we selected 19 clusters that manifested concrete changes. We discuss about our findings in Section \ref{}.

>>>>

An ASTDiff may include an arbitrary number of semantic changes. We are interested in finding the most important change within an ASTDiff and encoding that change in our datapoints. Sections \ref{sec:path_extraction} and \ref{sec:weighting_scheme} provide a detailed explanation of how we select the most important semantic information, which then helped us to obtain clusters that contained similar datapoints.

We used Pydriller, which is a repository miner for Python. Our target repositories were the top 18 most starred open source Rust projects on GitHub. We mined all the bug related commits and ran them through our pipeline. Then, we applied the DBSCAN clustering algorithm on the obtained datapoints. In Section \ref{sec:clustering_data} we discuss why we chose DBSCAN for clustering and how we tuned it to improve clusters quality.

\subsection{\label{sec:data_modelling}Data modelling}

\subsubsection{\label{sec:parsing_programs}Parsing Programs}

@@

\subsubsection{\label{sec:path_extraction}Path Extraction}

\begin{figure}[h]
\centering
\includegraphics[width=0.8\textwidth]{figs/extraction.png}
\caption{\label{fig:extraction}Path Extraction}
\end{figure}

@@

\subsubsection{\label{sec:weighting_scheme}Weighting Scheme}

@@

\subsection{\label{sec:mining_repositories}Mining Repositories}

\begin{figure}[h]
\centering
\includegraphics[width=1\textwidth]{figs/repos.png}
\caption{\label{table:repos} Target Repositories}
\end{figure}

@@

\subsection{\label{sec:clustering_data}Clustering Data}

@@

\begin{figure}[h]
\centering
\includegraphics[width=1\textwidth]{figs/clusters.png}
\caption{\label{fig:clustering} Clustering Experiments}
\end{figure}

@@

\subsubsection{\label{sec:manual_analysis_parameter_tuning}Manual Analysis for parameter tuning}

@@

\subsubsection{\label{sec:manual_analysis_cluster_selection}Manual Analysis for cluster selection}
% 3 types, refactoring dropped

@@

In our manual analysis, we randomly picked 50 datapoints of each cluster. If the cluster had less than 50 datapoints, we would analyze all of the datapoints in the cluster. Next, we read the code and also the bug report related to all the datapoints (if existed). In case that a similar pattern was detected over the instances of a cluster, we would write a natural language description of that cluster. After providing all the descriptions, we linked similar clusters together as possible candidates for merging. At last, after accomplishing re-examination for merge possibilities, we selected 19 clusters that manifested concrete changes. We discuss about our findings in Section \ref{sec:common_patterns} and \ref{sec:bc_patterns}.

### 2022-09-20 sections/results.tex
warnings fixed

\subsection{Common Bug Fix Patterns (RQ2)}

@@

\noindent \textit{ \textbf{G.attr.struct.} Modifying the attributes of structs}

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
// G.attr.struct.add: Adding attributes to a struct
++ #[derive(Debug, Clone)]
pub struct Call {
/// body 
}

@@

\noindent \textbf{Description:} In this pattern, the attribute set of a struct changes. This modification can happen in three different ways; (G1.1) Adding attributes to the struct, (G1.2) Removing attributes from the struct, (G1.3) Changing the content of an existing attribute. The attributes are Outer Attributes as we can find struct related non-terminals within the ASTDiff of the instances of this pattern. \\

@@

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
// G.attr.struct.field.add: Adding attributes to struct fields
#[derive(Serialize)]
pub struct Retain {
    ++ #[serde(skip_serializing_if = "is_empty")]
    pub attributes: Option<Attributes>,
}

@@

\noindent \textbf{Description:} In this pattern, the attribute set of a struct field changes. As with G1, this modification also happens in three different ways. Using attributes for struct fields, apart from being able to store different meta-information for different struct fields, for some use cases, the programmers can control the action they want to apply on different fields.

@@

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
// G.struct.field.add: Adding a new field to a struct
struct InnerListeners {
    pending: Mutex<Vec<Pending>>,
    <@\textcolor{red}{--}@> queue_object_name: Uuid,
}

@@

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]

@@

\noindent \textit{ \textbf{G.type.field} Changing a struct field type}

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
pub struct Manifest {
  -- pub features: Vec<String>,
  ++ pub features: HashSet<String>,
}

@@

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
#[derive(Debug, Clone, PartialEq)]
pub struct Anchor {
    -- point: Point<usize>,
    ++ point: Point<isize>,
    side: Side,
}

@@

\noindent \textbf{Description:} This a change of the generic (parametric) type of a struct field. Similar to G7, this change can also happen for the mentioned reasons. \\

@@

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
#[derive(Debug, Clone, PartialEq, Eq)]
pub enum ResolvedDependency {
  Resolved(ModuleSpecifier),
  -- Err(String),
  ++ Err(ResolvedDependencyErr),
}

@@

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
#[allow(dead_code)]
pub struct CoreState {
    // Other fields
    -- pending_views: Vec<ViewId>,
    ++ pending_views: Vec<(ViewId, Table)>,
    peer: Client
}
\end{lstlisting}

@@

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-- pub struct EventLoop<T: tty::EventedReadWrite> {
++ pub struct EventLoop<T: tty::EventedPty> {
    poll: mio::Poll,
    // Other fields
}

@@

\subsection{BC-Related Bug Fix Patterns (RQ3)}

@@

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-- let field_id = schema.get_or_create(attribute.clone())?;
++ let field_id = schema.get_or_create(&attribute)?;
\end{lstlisting}

@@

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
start_plugin_process(
--  manifest.clone(),
++  manifest,
    self.next_plugin_id(),
    self.self_ref.as_ref().unwrap().clone(),
);
\end{lstlisting}

@@

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
let repo = replace(&mut *contents, Processing).inner_repo();
-- let statuses = repo_to_statuses(repo, &self.workdir);
++ let statuses = repo_to_statuses(&repo, &self.workdir);
\end{lstlisting}

@@

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
enums.iter().for_each(|e| {
    let mut enum_template = EnumTemplate::new();
    -- enum_template.set_message_enum(&e);
    ++ enum_template.set_message_enum(e);
    // Other statements
});
\end{lstlisting}

@@

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
for p in wixobjs {
--  args.push(p.to_string());
++  args.push((*p).to_string());
}

@@

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
parser::Parser::new(args)
        .parse_module()
--      .map_err(|e| {
++      .map_err(|mut e| {
            e.emit();
            ()
        })?
\end{lstlisting}

@@

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-- let mut tx = tx.clone();
++ let tx = tx.clone();
\end{lstlisting}

@@

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
pub fn expand_delimited_square(
--  children: &Vec<TokenNode>,
++  children: &[TokenNode],
) -> Result<hir::Expression, ParseError> {
    // body
}
\end{lstlisting}
\noindent \textbf{Description:} In struct field definition or the types of the function formal parameters, type \verb+&Vec<T>+ changes to \verb+&[T]+. Clippy can detect the opportunity to apply this pattern. The reason is reducing type complication, as slice types in the form of \verb+&[T]+ or \verb+&str+ are usually sufficient for most use cases.

@@

\begin{lstlisting}[language=Rust, style=colouredRust, label={l3}]
-- const QUEUE_SIZES: &'static [u16] = &[QUEUE_SIZE];
++ const QUEUE_SIZES: &[u16] = &[QUEUE_SIZE];

>>>>

\subsection{\label{sec:common_patterns}Common Bug Fix Patterns (RQ2)}

@@

\noindent \textit{\label{sec:G.attr.struct}\textbf{G.attr.struct} Modifying the attributes of structs}

\begin{lstlisting}[language=Rust, style=colouredRust]
// G.attr.struct.add: Adding attributes to a struct
++ #[derive(Debug, Clone)]
pub struct Call {
/// body 
}

@@

\noindent \textbf{Description:} In this pattern, the attribute set of a struct changes. This modification can happen in three different ways; (G.attr.struct.add) Adding attributes to the struct, (G.attr.struct.drop) Removing attributes from the struct, (G.attr.struct.change) Changing the content of an existing attribute. The attributes are Outer Attributes as we can find struct related non-terminals within the ASTDiff of the instances of this pattern. \\

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
// G.attr.struct.field.add: Adding attributes to struct fields
#[derive(Serialize)]
pub struct Retain {
    ++ #[serde(skip_serializing_if = "is_empty")]
    pub attributes: Option<Attributes>,
}

@@

\noindent \textbf{Description:} In this pattern, the attribute set of a struct field changes. As with \textit{\textbf{G.attr.struct}}, this modification also happens in three different ways. Using attributes for struct fields, apart from being able to store different meta-information for different struct fields, for some use cases, the programmers can control the action they want to apply on different fields.

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
// G.struct.field.add: Adding a new field to a struct
struct InnerListeners {
    pending: Mutex<Vec<Pending>>,
    <@\textcolor{red}{--}@> queue_object_name: Uuid,
}

@@

\begin{lstlisting}[language=Rust, style=colouredRust]

@@

\noindent \textit{\label{sec:G.type.field}\textbf{G.type.field} Changing a struct field type}

\begin{lstlisting}[language=Rust, style=colouredRust]
pub struct Manifest {
  -- pub features: Vec<String>,
  ++ pub features: HashSet<String>,
}

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
#[derive(Debug, Clone, PartialEq)]
pub struct Anchor {
    -- point: Point<usize>,
    ++ point: Point<isize>,
    side: Side,
}

@@

\noindent \textbf{Description:} This a change of the generic (parametric) type of a struct field. Similar to \textit{\textbf{G.type.field}}, this change can also happen for the mentioned reasons. \\

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
#[derive(Debug, Clone, PartialEq, Eq)]
pub enum ResolvedDependency {
  Resolved(ModuleSpecifier),
  -- Err(String),
  ++ Err(ResolvedDependencyErr),
}

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
#[allow(dead_code)]
pub struct CoreState {
    // Other fields
    -- pending_views: Vec<ViewId>,
    ++ pending_views: Vec<(ViewId, Table)>,
    peer: Client
}
\end{lstlisting}

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
-- pub struct EventLoop<T: tty::EventedReadWrite> {
++ pub struct EventLoop<T: tty::EventedPty> {
    poll: mio::Poll,
    // Other fields
}

@@

\subsection{\label{sec:bc_patterns}BC-Related Bug Fix Patterns (RQ3)}

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
-- let field_id = schema.get_or_create(attribute.clone())?;
++ let field_id = schema.get_or_create(&attribute)?;
\end{lstlisting}

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
start_plugin_process(
--  manifest.clone(),
++  manifest,
    self.next_plugin_id(),
    self.self_ref.as_ref().unwrap().clone(),
);
\end{lstlisting}

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
let repo = replace(&mut *contents, Processing).inner_repo();
-- let statuses = repo_to_statuses(repo, &self.workdir);
++ let statuses = repo_to_statuses(&repo, &self.workdir);
\end{lstlisting}

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
enums.iter().for_each(|e| {
    let mut enum_template = EnumTemplate::new();
    -- enum_template.set_message_enum(&e);
    ++ enum_template.set_message_enum(e);
    // Other statements
});
\end{lstlisting}

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
for p in wixobjs {
--  args.push(p.to_string());
++  args.push((*p).to_string());
}

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
parser::Parser::new(args)
        .parse_module()
--      .map_err(|e| {
++      .map_err(|mut e| {
            e.emit();
            ()
        })?
\end{lstlisting}

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
-- let mut tx = tx.clone();
++ let tx = tx.clone();
\end{lstlisting}

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
pub fn expand_delimited_square(
--  children: &Vec<TokenNode>,
++  children: &[TokenNode],
) -> Result<hir::Expression, ParseError> {
    // body
}
\end{lstlisting}
\noindent \textbf{Description:} In struct field definition or the types of the function formal parameters, type \verb+&Vec<T>+ changes to \verb+&[T]+. Clippy can detect the opportunity to apply this pattern. The reason is reducing type complication, as slice types in the form of \verb+&[T]+ or \verb+&str+ are usually sufficient for most use cases.

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
-- const QUEUE_SIZES: &'static [u16] = &[QUEUE_SIZE];
++ const QUEUE_SIZES: &[u16] = &[QUEUE_SIZE];
